Exploring dimensions of biodiversity in Japanese ferns

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
update slides

---
background-position: 40% 50%
background-image: url(images/japan_intro.png)
background-size: contain

## A (very brief) introduction to the biogeography of Japan

.pull-left[
- Wide variation in climate from north (subarctic) to south (subtropical)

@@

@@

```{r richness, out.width = "130%"}
# Fern richness
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "No. taxa") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  ) +
  ylim(ja_lat_min, ja_lat_max)

@@

.footnote[Trendline: General additive model (GAM) with cubic splines]
---
## Other raw values are correlated with richness

@@

b <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(
    title = "Phylogenetic diversity",
    subtitle = "Faith's PD") +
  simple_map_theme()

c <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(
    title = "Functional diversity",
    subtitle = "Faith's PD measured\non trait dendrogram") +
  simple_map_theme()

@@

---
## Other raw values are correlated with richness... as expected

@@

.footnote[Trendline: GAM with cubic splines]
---
## Compare branch lengths with <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversity

RPD = original branch lengths / all branch lengths equal

@@

.footnote[Trendline: GAM with cubic splines]
---
## <u>**R**</u>elative <u>**F**</u>unctional <u>**D**</u>iversity: functional analog of RPD

@@

.middle[
<img src="images/daru_2017.png" alt="drawing" height="400"/>
]

@@

```{r tax-bioregion, out.width = "120%", fig.asp = 0.7}
ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp_light() +
  labs(
    title = "Taxonomic bioregions"
  )
```

@@

.footnote[`r Citet(bib, "shimizu_process_2014")`]

@@

```{r protected-areas, out.width = "120%", fig.asp = 0.7, fig.retina=200}
japan_box <- sf::st_bbox(ses_div_ferns_spatial)

@@

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE}
PrintBibliography(bib, start = 1, end = 5)
```
]

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE}
PrintBibliography(bib, start = 6, end = 10)
```
]

>>>>

---
## A (very brief) introduction to the biogeography of Japan

.pull-left-narrow[
- Wide variation in climate from north (subarctic) to south (subtropical)

@@

.pull-right-wide[
<img src="images/japan_intro_cropped.png" alt="drawing" height="580"/>
]

@@

```{r richness, out.width = "120%"}
# Fern richness
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "No. taxa") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  ) +
  ylim(ja_lat_min, ja_lat_max)

@@

.footnote[Trendline: General additive model (GAM) with cubic splines]

---
class: big-margin
## Measuring diversity with a tree

.center[
<img src="images/pd.png" alt="drawing" height="250"/>
]

<u>**P**</u>hylogenetic <u>**D**</u>iversity: Total unique branch length `r AutoCite(bib, "Faith1992")`

<u>**F**</u>unctional <u>**D**</u>iversity: PD, but measured with a trait dendrogram `r AutoCite(bib, "Petchey2002")`

---
## PD and FD are correlated with richness

@@

b <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(
    title = "Phylogenetic diversity") +
  simple_map_theme()

c <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(
    title = "Functional diversity") +
  simple_map_theme()

@@

---
## PD and FD are correlated with richness... as expected

@@

.footnote[Trendline: GAM with cubic splines]
---
background-position: 40% 50%
background-image: url(images/rpd.png)
background-size: contain

## Compare branch lengths with <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversity

.pull-right[
Long branches

- Refugium

- Occurrence of a few members of clades that mainly occur outside of the study region

Short branches

- Recently diverged lineages

.small[`r Citet(bib, "Mishler2014")`]
]

@@

.footnote[Trendline: GAM with cubic splines `r NoCite(bib, "Faith1992")`]
---
## <u>**R**</u>elative <u>**F**</u>unctional <u>**D**</u>iversity: functional analog of RPD

@@

.center[
<img src="images/daru_2017.png" alt="drawing" height="400"/>
]

@@

```{r tax-bioregion, out.width = "120%", fig.asp = 0.6}
ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp_light() +
  labs(
    title = "Taxonomic bioregions"
  )
```

@@

.footnote[`r Citet(bib, "Shimizu2014")`]

@@

```{r protected-areas, out.width = "120%", fig.asp = 0.65, fig.retina=200}
japan_box <- sf::st_bbox(ses_div_ferns_spatial)

@@

---
class: big-margin
## A growing threat: herbivory by deer (*Cervus nippon*)

.center[
<img src="images/deer_damage_1.png" alt="drawing" height="450"/>
]

---
class: big-margin
## A growing threat: herbivory by deer (*Cervus nippon*)

.center[
<img src="images/deer_damage_2.png" alt="drawing" height="500"/>
]

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
PrintBibliography(bib, start = 1, end = 5)
```
]

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
PrintBibliography(bib, start = 6, end = 10)
```
]
